---
title: "Kursusgang 7"
author: "Jonathan Strandberg"
date: "13 mar 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(astsa)
library(fracdiff)
library(forecast)
```

## Opgave 5.1 [ShSt]
Datasættet arf er 1000 simulerede observationer fra en ARFIMA$(1,1,0)$ model med $\phi=0.75$ og $d=0.4$
a) Plot the data and comment
```{r}
plot(arf , main = "Simuleret ARFIMA(1,1,0)")
```

Ser ikke stationært ud

b) Plot the ACF and PACF of the data and comment
```{r}
acf(arf) ; pacf(arf)
```

PACF "cutter off" efter lag 2, og tyder på at det er en AR(2) model

c) Estimate the parameters and test for the significance of the estimates $\widehat{\phi}$ and $\widehat{d}$
```{r}
summary(fracdiff(arf , nar = 1))
```
Ses at både $\widehat{\phi}$ and $\widehat{d}$ er meget signifikante

d) Explain why, using the results of parts (a) and (b), it would seem reasonable to difference the data prior to the analysis. That is, if $x_t$ represents the data, explain why we might choose to fit an ARMA model to $\Delta x_t$

Ser ikke stationæer ud så det er rimeligt at diffe

e) Plot the ACF and PACF of $\Delta x_t$ and comment
```{r}
darf <- diff(arf)

acf(darf) ; pacf(darf)
```

Ved at se på ACF ligner det at $\Delta x_t$ er stationær

f) Fit an ARMA model to $\Delta x_t$ and comment
```{r}
auto.arima(darf)
sarima(darf , 1 , 0 , 0 , details = FALSE )$ttable
```

