---
title: "Selvstudie 2"
author: "Jonathan Strandberg"
date: "18 mar 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stats)
library(astsa)
library(forecast)
library(tseries)
```

## Problem 10.2
An AR model has AR characteristic polynomial
\[
(1 - 1.6x + 0.7x^2)(1 - 0.8x^{12})
\]

a) Is the model stationary?

\[
1 - 1.6x + 0.7x^2 -0.8x^{12} + 1.28x^{13} - 0.56x^{14}
\]
```{r}
coeff <- c(1,-1.6,0.7,0,0,0,0,0,0,0,0,0,-0.8,1.28,-0.56)
roots <- polyroot(coeff)
sapply(roots , norm , type="2")

# arima.sim(list(ar = coeff[-1]),n = 1000) # Kan ikke køre ('ar' part of model is not stationary)
```

Ingen rødder på enhedscirklen $\Leftrightarrow$ stationær


b) Identify the model as certain seasonal ARIMA model.


## Problem 10.7
Suppose that the process $\{Y_t\}$ develops according to $Y_t = Y_{t-4} + e_t$ with $Y_t = e_t$ for $t = 1,2,3,4$

a) Find the variance function for $\{Y_t\}$
\[
  Var[Y_t] = ceiling\left(\frac{t}{4}\right)^2 \sigma_e^2
\]

b) Find the autocorrelation function for $\{Y_t\}$

c) Identify the model for $\{Y_t\}$ as a certain seasonal ARIMA